\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{amsmath, amssymb, graphicx, epsfig, fleqn}
\setlength{\parindent}{0pt}
\newcommand{\ud}{\,\mathrm{d}}
\everymath{\displaystyle}
\def\FillInBlank{\rule{2.5in}{.01in} }
\pagestyle{empty}

\begin{document}
\begin{center}
\Large
\rm{Math 111}
\\
\rm{Exercise answers:  3.8 \& 3.9}
\\
\end{center}
\vspace{0.2in}
\fboxsep0.5cm


{\bf 3.8}

\begin{enumerate}
\item{It is determined that an invasive species of fish grows with a constant relative growth rate and that the population doubles in 3 years
\begin{enumerate}
\item{How long does it take to increase by a factor of 10? (9.97 years) }
\item{What is the value of the relative growth rate?  (0.231 fish/year/fish)}
\end{enumerate}
}

\item{A certain radioactive material has half-life of 2.5 years. 
\begin{enumerate}
\item{How long until a mass of 300 grams of this material decays to 5 grams?  (14.77 years)}
\item{What is the decay rate at the time it reaches mass 5 grams? (-1.39  gram/year)}
\end{enumerate}
}

\item{Let $v(t)$ be the velocity of a falling object of mass 1 kg.  A simple model for the velocity is the equation $v' = 9.8 - 0.1v$.
  \\ (In this model, 9.8 is the acceleration due to gravity, and $0.1v$ is the drag force that the object experiences as it falls.)
\begin{enumerate}
\item{Find a formula for  $v(t)$ if $v(0)=0$. ($v(t) = 98(1-e^{-0.1t})$) }
\item{Determine what happens to the velocity as $t\to\infty$.} ($v \to 98$.  This is the terminal velocity.)
\end{enumerate}

}
\end{enumerate}


{\bf 3.9}

\begin{enumerate}
  \item{A cylindrical tank of radius 5m is being drained at a rate of 3m$^3$/min.  How fast is the depth of the water in the tank decreasing
  when the depth is 7m?  (-3/25$\pi$)}
  

\item{Police radar is stationed near a road.  The officer aims the radar gun at a passing car and the gun is at an angle of 45$^{\circ}$ to the road
  If the officer records the distance between the car and the radar is decreasing at a rate of 100 km/hr, how fast is the car travelling? (141.42 kph)
}

\item{Sand is dumped on a pile in the shape of a cone at a rate of 30 ft$^3$/min.  The diameter and the height of the cone remain equal as the pile
  grows.  How fast is the height of the pile increasing when the height is 10 ft?  (6/5$\pi$)
}

\item{Two cars start off from the same point.  One travels west at 25 mph, the other south at 60mph.  At what speed are they moving apart 2 hours later?  (65 mph)
}
\item{In a certain species of fish, brain weight (in grams) as a function of body weight (in grams) is determied to be $B=0.007W^{2/3}$.  Body weight as a function of body length is determined to be $W=0.12L^{5/2}$.  How fast is the fish brain growing when the fish is 5 cm and growing by 0.1cm/day?  ($8.3\times10^{-4}$ grams/day)
}

\item{If the radius of a cylinder is decreasing by 2cm/min, and the height is increasing at a rate of 10 cm/min, how is the volume of the cylinder changing when the radius is 5 cm and the height is 15 cm?  ($-50\pi$)}

\item{The area between two varying concentric circles is at all times $9\pi$ in$^2$. If the area of the larger circle is increasing by 10$\pi$ in$^2$/sec, how
fast is the circumference of the smaller circle changing?  (1.25 in/sec)}

\item{A swimming pool is 12 m long, 7 m wide, 1.2 m deep at the shallow end and 3 m deep at the deep end.  Water is being pumped into the pool
  at a rate of $\frac13$m$^3$/min.  How fast is the water rising when the water is 1 m deep at the deep end?
(1/140 m/min)}

\end{enumerate}



\end{document}


