\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{amsmath, amssymb, graphicx, epsfig, fleqn}
\setlength{\parindent}{0pt}
\newcommand{\ud}{\,\mathrm{d}}
\everymath{\displaystyle}
\def\FillInBlank{\rule{2.5in}{.01in} }
\pagestyle{empty}

\begin{document}
\begin{center}
\Large
\rm{Math 111}
\\
\rm{Chapter 2.7:  Derivatives \& Rates of Change}
\\
\end{center}
\vspace{0.2in}
\fboxsep0.5cm

(DEFINITION) The {\bf derivative of a function $f$ at a number $a$} is:
\begin{displaymath}
f'(a) = \lim_{h\to 0} \frac{f(a+h)-f(a)}{h}
  \end{displaymath}


\vspace{0.5in}

The meaning of $f'(a)$ is
\begin{enumerate}
\item{}
  \vspace{0.2in}
\item{}
  \vspace{0.2in}
  
  \end{enumerate}

  \vspace{0.2in}

(EXAMPLE)  Find $f'(9)$ if $f(x) = 4-\sqrt{x}$.  Sketch the graph of $f$ and the tangent line at $x=9$.


\pagebreak


(EXAMPLE)  The height of a falling object in meters is given by $y(t) = 150 - 4.9t^2$, where time $t$ is measured
in seconds.  Find the velocity of the object at the time $t=5$.  What are the units of $y'(5)$?


\vspace{3.5in}


(EXAMPLE)  Find the derivative $p'(2)$ if $p(x) = 3/x$.  Sketch the graph of $p$, the line tangent to the
graph at $(2, 1.5)$, and give the equation of the line.


\pagebreak

(INTERPRETATIONS)\\

\vspace{0.1in}

In certain circumstances, the volume of a gas is inversely proportional to the pressure.  We could write
volume as a function of pressure $V(P) = k/P$ where $k$ is a constant, $V$ is measured in cm$^3$ and pressure measured
in kPa.
\begin{enumerate}
\item{What does $V'(100)$ measure?}
\item{What is the sign of $V'(100)$ ?}
\item{What are the units $V'(100)$ ?}
\item{Which do you think is larger, $V'(100)$ or $V'(500)$ ?}
\end{enumerate}

\vspace{0.2in}


The garbage $G$ produced by a city is a function of its population $P$.  Suppose $P$ is measured in thousands and
$G$ is measured in tons.  
\begin{enumerate}
\item{What does $G'(250)$ measure?}
\item{What is the sign of $G'(250)$ ?}
\item{What are the units $G'(250)$ ?}
\end{enumerate}


\vspace{0.2in}


Let $A(t)$ represent the total mass of apples (in kg) you gather as a function of time you spend in an orchard (in hours).
\begin{enumerate}
\item{What is the sign of $A'(6)$ ?}
\item{What are the units $A'(6)$ ?}
\item{Which do you think is larger $A'(6)$ or $A'(200)$?  Why? }
\end{enumerate}


\vspace{0.2in}

Suppose we know that $g$ is a continuous function with $g(1) = 2$, $g'(1) = 0$, $g(4) = 0$, and $g'(4) = 1 $.
Sketch a possible graph of $g$.




\end{document}


